Portfolio prioritization of salmon conservation

### 2013-04-09 inst/examples/salmonport.Rmd
at copy sent to committee before 2013-04-08

@@

```{r env-ts, fig.height=4, fig.width=4, cache = FALSE, fig.cap="Example temperature time series\\label{fig:env-ts}"}
#library(metafolio)
#or for rapid development:
metafolio_files <-  list.files(path = "../../R/")
for(i in 1:length(metafolio_files)){
  source(paste("../..//R/", metafolio_files[i], sep = ""));rm(i)}

source("plot_env_ts.R")
col_pal <- rev(gg_color_hue(10))
```

@@

\begin{table}[h!]
\centering
\small
\caption{Salmon metapopulation parameters with base case and alternate values.}
\begin{tabular}{p{7.0cm}p{1.6cm}p{3.2cm}}
\toprule
Description & Parameter & Base [lower, upper] \\
\midrule
Stock-recruit residual standard deviation (on log scale) & $\sigma_v$ & 0.30 [0.05, 0.50] \\
First order (AR1) serial correlation of stock-recruit residuals & $\rho_w$ & 0.40 [0, 0.80] \\
Fraction of fish that stray from natal streams & $f_{stray}$ & 0.02 [0, 0.10] \\
Exponential rate of decay of straying with distance & $m$ & 0.30 [0.05, 0.50] \\
Standard deviation of beta distribution for implementation error & $\sigma_{impl}$ & 0.05 [0, 0.20] \\
Frequency of assessment (years) & $f_{assess}$ & 20 [5, 50] \\
\bottomrule
\end{tabular}
\label{tab:salm-pars}
\end{table}

@@

```{r fig.width=6, fig.height=3.75, fig.cap="Example Ricker curves for the minimum-variance metapopulation."}
plot_rickers(min_var_sim, pal = col_pal)
```

@@

```{r plot-weights-median, fig.width=4.2, fig.height=3.2, cache = FALSE, fig.cap="Median (solid line) and 25th and 75th quantiles of the weights across 40 minimum-variance portfolios. Just another way of looking at it."}
source("min_var_ports_dist.R")
```

@@

```{r plot-cons-plans-mv, fig.width=6.5, fig.height=3.5, fig.cap="Conservation plans in mean-variance space. The dots show portfolios from individual simulation runs. The dark and light shaded areas show 25% and 75% quartiles. The left panel shows a scenario with a stationary environmental signal and strong short-term environmental noise. The right panel shows a scenario with a linearly-increasing environmental signal and minimal short-term environmental noise. These two scenarios bound the two environmental extremes: strong short-term noise with minimal long-term signal vs. a strong long-term signal with minimal short-term signal. Notice what happens to the two *half* conservation strategies. A more realistic scenario combines the two. The efficient frontier is shown for the displayed portfolios. This isn't necessarily the overall efficient frontier.", cache=FALSE}
#Notes #
#-b is "unfished equilibrium stock size"
require(RColorBrewer)
cols <- brewer.pal(5, "Dark2")[c(1,2,4,5,3)] 
source("cons_plans.R")
```

@@

# Take home messages
1. If the dominant environmental signal is a long-term shift, and you keep around one half of the response diversity, then you get different portfolio means but the same variances.
2. If the dominant environmental signal is short-term, and you keep around one half of the response diversity, then you get different portfolio variances but the same means.
3. Combine those two (which best reflects reality) and you get both different means and variances.
4. Whether you invest in just stable, just asynchronous, or a mix of both, all that really matters is that you have as many populations as possible and as balanced as possible response diversity.
5. An even weighting across as many possible streams (assuming balanced response diversity) gets you pretty close to the minimum-variance portfolio. [not shown currently]
6. With systematic long-term environmental change that dominates: if you can forecast the environment and know the responses of salmon populations from individual streams then you can increase your expected rate of metapopulation growth with minimal effect on variance by prioritizing the populations that will do well. Although if you get it wrong you will do worse than if you had hedged your bets and conserved response diversity.
7. But, if there are also strong short-term environmental fluctuations (and there probably are), then there's always a benefit to portfolio variance by having some response diversity.
8. Variable stocks aren't inherently undesirable to conserve, as long as they are balanced or averaged out.
9. Take home message: "To keep every cog and wheel is the first precaution of intelligent tinkering." But, if you must prioritize, keep a selection of different cogs and
 wheels --- especially a spatially diverse set of cogs and wheels. That gives you the best chance of an efficient ecological portfolio.

@@

>>>>

*Other title ideas:*  
- Portfolio optimization reveals...  
- something more general with animals instead of just salmon?  
- Portfolio theory reveals conservation strategies that...  
- Intelligent tinkering with ecological portfolios: patterns of environmental response diversity drive salmon conservation priorities  
- Keeping every cog and wheel...

@@

```{r env-ts, fig.height=4, fig.width=4, cache = FALSE, fig.cap="Example temperature time series\\label{fig:env-ts}"}
#library(metafolio)
#or for rapid development:
library(devtools)
load_all("../../") # assuming you're 2 folders deeper than the package
source("plot_env_ts.R")
col_pal <- rev(gg_color_hue(10))
```

@@

\begin{table}[h!]
\centering
\small
\caption{Salmon metapopulation parameters with base case and alternate values.}
\begin{tabular}{p{7.0cm}p{1.6cm}p{3.2cm}}
\toprule
Description & Parameter & Base [lower, upper] \\
\midrule
Stock-recruit residual standard deviation (on log scale) & $\sigma_v$ & 0.30 [0.05, 0.50] \\
First order (AR1) serial correlation of stock-recruit residuals & $\rho_w$ & 0.40 [0, 0.80] \\
Fraction of fish that stray from natal streams & $f_{stray}$ & 0.02 [0, 0.10] \\
Exponential rate of decay of straying with distance & $m$ & 0.3 [0.05, 0.5] \\
Standard deviation of beta distribution for implementation error & $\sigma_{impl}$ & 0.05 [0, 0.20] \\
Frequency of assessment (years) & $f_{assess}$ & 20 [5, 50] \\
\bottomrule
\end{tabular}
\label{tab:salm-pars}
\end{table}

@@

```{r fig.width=6, fig.height=3.75, fig.cap="Example Ricker curves for the minimum-variance metapopulation. (Ignore right now)"}
plot_rickers(min_var_sim, pal = col_pal)
```

@@

```{r plot-weights-median, fig.width=4.2, fig.height=3.2, cache = FALSE, fig.cap="Median (solid line) and 25th and 75th quantiles of the weights across 40 minimum-variance portfolios. Just another way of looking at it. (Nothing to see here right now.)"}
source("min_var_ports_dist.R")
```

@@

```{r plot-cons-plans-mv, fig.width=6.5, fig.height=3.5, fig.cap="Conservation plans in mean-variance space. The dots show portfolios from individual simulation runs. The dark and light shaded areas show 25% and 75% quartiles. The left panel shows a scenario with a stationary environmental signal and strong short-term environmental noise. The right panel shows a scenario with a linearly-increasing environmental signal and minimal short-term environmental noise. These two scenarios bound the two environmental extremes: strong short-term noise with minimal long-term signal vs. a strong long-term signal with minimal short-term signal. Notice what happens to the two *half* conservation strategies. A more realistic scenario combines the two kinds of environmental noise. The efficient frontier is shown for the displayed portfolios. This isn't necessarily the overall efficient frontier.", cache=FALSE}
#Notes #
#-b is "unfished equilibrium stock size"
require(RColorBrewer)
cols <- brewer.pal(5, "Dark2")[c(1,2,4,5,3)] 
source("cons_plans.R")
```

@@

# Some thoughts on the results
1. If the dominant environmental signal is a long-term shift, and you keep around one half of the response diversity, then you get different portfolio means but the same variances.
2. If the dominant environmental signal is short-term, and you keep around one half of the response diversity, then you get different portfolio variances but the same means.
3. Combine those two (which best reflects reality) and you get both different means and variances.
4. Whether you invest in just stable, just asynchronous, or a mix of both, all that really matters is that you have as many populations as possible and as balanced as possible response diversity.
5. An even weighting across as many possible streams (assuming balanced response diversity) gets you pretty close to the minimum-variance portfolio. [not shown currently]
6. With systematic long-term environmental change that dominates: if you can forecast the environment and know the responses of salmon populations from individual streams then you can increase your expected rate of metapopulation growth with minimal effect on variance by prioritizing the populations that will do well. Although if you get it wrong you will do worse than if you had hedged your bets and conserved response diversity.
7. But, if there are also strong short-term environmental fluctuations (and there probably are), then there's always a benefit to portfolio variance by having some response diversity.
8. Variable stocks aren't inherently undesirable to conserve, as long as they are balanced or averaged out.
9. Take home message: "To keep every cog and wheel is the first precaution of intelligent tinkering." But, if you must prioritize, keep a selection of different cogs and
 wheels --- especially a spatially diverse set of cogs and wheels. That gives you the best chance of an efficient ecological portfolio.

@@

\includepdf{salmonport-mindmap.pdf}
